---
title: "PCA - Wine"
author: "Evelyn"
date: "1/22/2021"
output: html_document
toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Summary

PCA is a data dimension reduction method, to uncover latent variables. 
Not many of the variables in high-dimentionsal data are required. Some are highly correlated with others and these variables may be omitted, while retaining the same level of information in the dataset in terms of explaining the variance. 

It is used as an exploratory data analysis tool, and may be used for feature engineering and/or clustering. 

# Workflow

1. Import data
2. Exploratory data analysis

  - skim
  - ggcorr
  - ggpairs
  
3. Check assumptions on whether PCA can be carried out
  
  - KMO
  - Bartlett
  
4. Carry out PCA using tidymodels workflow

Always use only continuous variables, ensure that there are no missing data. 
Determine the number of components using eigenvalues, scree plots and parallel analysis. 
  
  - recipe : preprocess the data
  - prep : evaluate the data
  - bake : get the PCA Scores results
  - visualize
  - communicate results
  
The scores plot show the **positions** of the individual wine samples *in the coordinate system of the PCs*.

The loadings plot shows the **contribution of the X variables** to the PCs.

## Loading packages

```{r}
library(ChemometricsWithR)

library(pacman)
p_load(corrr, palmerpenguins, GGally, tidymodels, tidytext, tidyverse, psych,
       skimr, gridExtra, kohonen, janitor, learntidymodels, kohonen)
```

## Import

```{r}
data(wines)

wines <- as.data.frame(wines) %>% 
  janitor::clean_names() %>%  # require data.frame
  as_tibble() %>% 
  cbind(vintages)  # vintages = Y outcome = category
 
glimpse(wines)
```


## EDA

### skimr
```{r}
skim(wines) # 177 x 13, all numeric + Y outcome
```

```{r}
# types of wines
wines %>% 
  select(vintages) %>% 
  unique()  # Barolo, Grignolino, Barbera
```


### GGally

```{r}
wines %>% 
  select(-vintages) %>% 
  ggcorr(label = T, label_alpha = T, label_round = 2)

wines %>% 
  ggpairs(aes(col = vintages))
```


## Checking assumptions

```{r}
# Continuous Y
# No missing data
# Check assumptions for PCA #####
wines_no_y <- wines %>% 
  select(-vintages)

glimpse(wines_no_y)

# KMO
wines_no_y %>% 
  cor() %>% 
  KMO() # .70 above : YES

# Bartlett's test of sphericity
wines_no_y %>% 
  cor() %>% 
  cortest.bartlett(., n = 177) # p<0.05

```


## Tidymodels (PCA)

### recipe

```{r}
wines_recipe <- recipe(~ ., data = wines) %>% 
  update_role(vintages, new_role = "id") %>% 
  # step_naomit(all_predictors()) %>% 
  step_normalize(all_predictors()) %>% 
  step_pca(all_predictors(), id = "pca")

wines_recipe # 13 predictors

```


### prep

```{r}
wines_prep <- prep(wines_recipe)

wines_prep # trained

tidy_pca_loadings <- wines_prep %>% 
  tidy(id = "pca")

tidy_pca_loadings # values here are the loading
```


### bake

```{r}
wines_bake <- bake(wines_prep, wines)
wines_bake  # has the PCA SCORES in normal form
```


### check number of PC

```{r}

# a. Eigenvalues
# data is stored in penguins_prep, step 3

wines_prep$steps[[2]]$res$sdev # 3

# b. Scree plot/Variance plot

wines_prep %>% 
  tidy(id = "pca", type = "variance") %>% 
  filter(terms ==  "percent variance") %>% 
  ggplot(aes(x = component, y = value)) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  scale_x_continuous(breaks = 1:4) +
  labs(title = "% Variance explained",
       y = "% total variance",
       x = "PC",
       caption = "Source: ChemometricswithR book") +
  theme(axis.title = element_text(face = "bold", size = 12),
        axis.text = element_text(size = 10),
        plot.title = element_text(size = 14, face = "bold"))  # 2 or 3

# bii: Cumulative variance plot

wines_prep %>% 
  tidy(id = "pca", type = "variance") %>% 
  filter(terms == "cumulative percent variance") %>%
  ggplot(aes(component, value)) +
  geom_col(fill= "forestgreen") +
  labs(x = "Principal Components", 
       y = "Cumulative variance explained (%)",
       title = "Cumulative Variance explained") +
  geom_text(aes(label = round(value, 2)), vjust = -0.2, size = 4) +
  theme_minimal() +
  theme(axis.title = element_text(face = "bold", size = 12),
        axis.text = element_text(size = 10),
        plot.title = element_text(size = 14, face = "bold")) 

# c. Parallel analysis

fa.parallel(cor(wines_no_y),
            n.obs = 333,
            cor = "cor",
            plot = T)  # 3

```



### visualize

#### Loadings plot

```{r}
plot_loadings <- tidy_pca %>% 
  filter(component %in% c("PC1", "PC2", "PC3", "PC4")) %>% 
  mutate(terms = tidytext::reorder_within(terms, 
                                          abs(value), 
                                          component)) %>% 
  ggplot(aes(abs(value), terms, fill = value>0)) +
  geom_col() +
  facet_wrap( ~component, scales = "free_y") +
  scale_y_reordered() + # appends ___ and then the facet at the end of each string
  scale_fill_manual(values = c("deepskyblue4", "darkorange")) +
  labs( x = "absolute value of contribution",
        y = NULL,
        fill = "Positive?",
        title = "PCA Loadings Plot",
        subtitle = "Number of PC should be 3, compare the pos and the neg",
        caption = "Source: ChemometricswithR") +
  theme_minimal()


plot_loadings
# PC1: flavonoids, tot_phenols, od_ratio, proanthocyanidins, col_hue, 36%
# PC2: col_int, alcohol, proline, ash, magnesium; 19.2%
# PC3: ash, ash_alkalinity, non_flav phenols; 11.2%
# PC4: malic acid?


```

#### Loadings only

```{r}
pca_loadings_only <- pca_loadings_wider %>% 
  ggplot(aes(x = PC1, y = PC2)) +
  geom_segment(aes(xend = PC1, yend = PC2),
               x = 0, 
               y = 0,
               arrow = arrow_style) +
  ggrepel::geom_text_repel(aes(x = PC1, y = PC2, label = terms),
            hjust = 0, 
            vjust = 1,
            size = 5,
            color = "red") +
  labs(title = "Loadings on PCs 1 and 2 for normalized data") +
  theme_classic()
  

pca_loadings_only
```


#### Scores plot
```{r}
# get pca loadings into wider format
pca_loadings_wider <- tidy_pca %>% 
  pivot_wider(names_from = component, id_cols = terms)

# Scores plot #####
# PCA SCORES are in bake
pc1pc2_scores_plot <- wines_bake %>% 
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = vintages, shape = vintages), 
             alpha = 0.8, size = 2) +
  scale_color_manual(values = c("deepskyblue4", "darkorange", "purple")) +
  labs(title = "Scores on PCs 1 and 2 for normalized data",
       x = "PC1 (36%)",
       y = "PC2 (19.2%)") +
  theme_classic() +
  theme(legend.position = "bottom") 

pc1pc2_scores_plot 
```

#### Scores + Loadings plot

```{r}
# does not work cos scale is different

# define arrow style 
arrow_style <- arrow(angle = 30,
                     length = unit(0.2, "inches"),
                     type = "closed")

# combined plot (does not work here) ######
pc1pc2_scores_and_loadings <- pc1pc2_scores_plot + 
  geom_segment(data = pca_loadings_wider,
               aes(xend = PC1, yend = PC2),
               x = 0, 
               y = 0,
               arrow = arrow_style) +
  geom_text(data = pca_loadings_wider,
            aes(x = PC1, y = PC2, label = terms),
            hjust = 0, 
            vjust = 1,
            size = 5,
            color = "forestgreen") +
  labs(title = "PC1 PC2 Scores + Loadings Plot") +
  theme(legend.position = "bottom")


pc1pc2_scores_and_loadings

```

#### Finalised plots
```{r}
grid.arrange(pc1pc2_scores_plot, pca_loadings_only, ncol = 2)
```


## Check against Data

```{r}
wines %>% 
  group_by(vintages) %>% 
  summarise(across(c(flavonoids, col_int, ash, malic_acid),
                   mean,
                   na.rm = T))
```


## Interpretation of results

PC1: flavoids, tot_phenols, OD_ratio ad proanth
PC2: col_intensity, alcohol, proline are determining factors
PC3: ash, ash_alkalinity are grouped together
PC4: malic acid

Barbera, indicated in blue, has the largest score on PC 1 and PC2.
Barolo, indicated in orange, has the smallest score on PC 1.
Grignolo, indicated in purple, has the lowest score on PC 2. 


Barbera has low flavonoids, high col_int and high malic acid
Barolo has high flavonoids, medium col_int and intermediate malic acid
Grignolino has intermediate flavonoids, high col_int and low malic acid.

# References

- https://rdrr.io/github/tidymodels/learntidymodels/f/inst/tutorials/pca_recipes/pca_recipes.Rmd
- https://allisonhorst.github.io/palmerpenguins/articles/articles/pca.html
