---
title: "Multiple t-test in R"
description: |
  What if I have many variables to compare?
author:
  - name: lruolin
date: 07-26-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 3
editor_options: 
  chunk_output_type: console
---

I was looking at a dateset at work, and was wondering how I can carry out t-test to check if there were any significant difference in flavor compounds between two different species of the same fruit. 

The exercise below is from a website <https://www.datanovia.com/en/blog/how-to-perform-multiple-t-test-in-r-for-different-variables/>, and I used it on a dataset which was on the physicochemical properties of bananas (from Easy Statistics for Food Science with R)


# Load Packages

```{r}
library(tidyverse)
library(rstatix)
library(ggpubr)
```


# Load dataset

```{r}
data <- iris %>% 
  filter(Species != "virginica") %>% 
  as_tibble()

glimpse(data)

```

# Transform into long format

```{r}

data_long <- data %>% 
  pivot_longer(cols = Sepal.Length:Petal.Width,
               names_to = "variables",
               values_to = "value")

glimpse(data_long)
```

# Run multiple t-test using rstatix package

```{r}
stat_test <- data_long %>% 
  group_by(variables) %>% 
  t_test(value ~ Species) %>% 
  adjust_pvalue(method = "bonferroni") %>% 
  add_significance()
```


# Visualization

```{r}
boxplot <- ggboxplot(data_long,
          x = "Species",
          y = "value",
          fill = "Species",
          palette = "jco",
          legend = "none",
          ggtheme = theme_pubr(border = T)) +
  facet_wrap ( ~ variables, scales = "free" )

# with sig diff
stat_test_plot <- stat_test %>% 
  add_xy_position(x = "Species") # from r-statix

boxplot + 
  stat_pvalue_manual(stat_test_plot, label = "p.adj.signif",
                     hide.ns = T, tip.length = 0)
```

# Applying it to a food science dataset

Comparing the pH and TSS (Brix) of bananas (Cavendish vs Dream)
(dataset is from Easy Statistics for Food Science with R, page 78)


## Dataset
```{r}

banana <- tribble(
  ~ variety,  ~pH,   ~tss,
  # --------#------#-------
  "cavendish", 5.60, 4.33,
  "cavendish", 5.57, 4.03,
  "cavendish", 4.76, 3.77,
  "cavendish", 5.56, 4.10,
  "cavendish", 4.95, 3.97,
  "cavendish", 4.84, 4.40,
  "cavendish", 5.07, 4.50,
  "cavendish", 4.94, 4.43,
  "cavendish", 5.04, 4.30,
  "cavendish", 4.93, 4.57,
  "cavendish", 5.05, 4.30,
  "cavendish", 5.21, 4.40,
  "dream", 4.31, 3.67,
  "dream", 4.41, 3.80,
  "dream", 4.35, 3.00,
  "dream", 4.49, 3.40,
  "dream", 4.39, 3.67,
  "dream", 4.43, 3.33,
  "dream", 4.44, 3.47,
  "dream", 4.44, 3.57,
  "dream", 4.52, 3.20,
  "dream", 4.79, 3.17,
  "dream", 4.68, 3.27,
  "dream", 4.83, 3.30
) %>% 
  mutate(variety = factor(variety))


glimpse(banana)
```

## Statistical test

```{r}
banana_long <- banana %>% 
  pivot_longer(cols = pH:tss,
               names_to = "variables",
               values_to = "values")

glimpse(banana_long)

stat_test <- banana_long %>% 
  group_by(variables) %>% 
  t_test(values ~ variety) %>% 
  adjust_pvalue(method = "none") %>% 
  add_significance()

stat_test
```

## Visualization

```{r}
boxplot <- ggboxplot(banana_long,
          x = "variety",
          y = "values",
          fill = "variety",
          palette = "jco",
          legend = "none",
          ggtheme = theme_pubr(border = T)) +
  facet_wrap ( ~ variables, scales = "free" ) +
  labs(title = "Cavendish Bananas have higher pH and total soluble solids. ")

# with sig diff
stat_test_plot <- stat_test %>% 
  add_xy_position(x = "variety") # from r-statix

boxplot + 
  stat_pvalue_manual(stat_test_plot, 
                     label = "p = {p.adj}{p.adj.signif}",
                     hide.ns = T, tip.length = 0)
```



# Reference:

<https://www.datanovia.com/en/blog/how-to-perform-multiple-t-test-in-r-for-different-variables/>

<https://www.datanovia.com/en/blog/how-to-add-p-values-to-ggplot-facets/>

<https://books.google.com.sg/books?id=e0JvDwAAQBAJ&pg=PA77&lpg=PA77&dq=cavendish+dream+pH+tss&source=bl&ots=XHNZARp1_F&sig=ACfU3U1QE4DIuf936cwJ5HdUQlhfGmtZNQ&hl=en&sa=X&ved=2ahUKEwjt46PMwf_xAhUEWCsKHVhwBBgQ6AEwEHoECBMQAw#v=onepage&q=cavendish%20dream%20pH%20tss&f=false>