---
title: "Date and Time in R"
description: |
  R4DS 13 - Lubridate
author:
  - name: lruolin
date: 05-24-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R4DS Practice 13: lubridate

The codes below are from the practice exercises in <https://r4ds.had.co.nz/>, and are taken with reference from: <https://jrnold.github.io/r4ds-exercise-solutions/>

# Let's begin now

Loading tidyverse package. 

```{r}
library(tidyverse)
library(lubridate)
library(nycflights13)
```

# Creating Date/Times from Strings

Identify the order in which the year, month and day appear in your dates, then arrange y, m , d in the same order. 

```{r}
ymd("2021-05-24")

# you can also use unquoted numbers

ymd(20210524)

# you can also add time using h, m, s

ymd_hms ("2021-05-24 20:11:09", tz = "Singapore")

# Use OlsonNames() to display timezone names

```

# Creating Dates from Individual Components

Sometimes, you have the individual components instead of teh date-time spread across multiple columns.

```{r}
flights %>% 
  select(year, month, day, hour, minute)

# To create date/time:
# use make_date()
# or make_datetime()

flights %>% 
  select(year, month, day, hour, minute) %>% 
  mutate(departure = make_datetime(year, month, day, hour, minute))
```
Use the appropriate lubridate function to parse each of the following dates:

```{r}
d1 <- "January 1, 2010"
mdy(d1)

d2 <- "2015-Mar-07"
ymd(d2)

d3 <- "06-Jun-2017"
dmy(d3)

d4 <- c("August 19 (2015)", "July 1 (2015)")
mdy(d4)

d5 <- c("12/30/14") # Dec 30, 2014
mdy(d5)


```




# Reference
<https://r4ds.had.co.nz/>

<https://jrnold.github.io/r4ds-exercise-solutions/>

