---
title: "Calibration Curves Data"
description: |
  R script for calculating Limit of Detection and Limit of Quantification
author:
  - name: lruolin
date: 01-15-2021
output:
  distill::distill_article:
    toc: true
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

# for the metadata above, can add in under author,name:
# url:, # affiliation, affiliation_url

```
*Note: the data and theory on calibration curve were with reference from: Statistics and Chemometrics for Analytical Chemistry, James N. Miller and Jane Charlotte Miller, 6th edition, Chapter 5*

# Loading required packages
```{r, echo = TRUE}
library(pacman)
p_load(tidyverse, broom, chemCal)
```


# Background
Chemists often work with calibration data using standards of known concentrations and putting them through instrumental analysis. When plotting a calibration curve, it is of interest to calculate the limit of detection (LOD) and limit of quantification (LOQ) of the method. 

# Import Example Dataset

The fluorescence intensities of standard aqueous fluorescein solutions were analysed with a spectrophotometer, and the fluorescence results are shown below:

```{r}
data <- tribble(
  ~conc_pgml, ~fluo,
  0,  2.1,
  2,  5.0,
  4,  9.0,
  6,  12.6,
  8,  17.3,
  10, 21.0,
  12, 24.7,
)

data
```

# Visualize
```{r}
theme_set(theme_bw())

data %>% 
  ggplot(aes(x = conc_pgml, y = fluo)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Calibration Plot of Fluorescence Units vs Fluorescein Concentration",
       x = "Concentration (pg/ml)",
       y = "Fluorescence Intensities") +
  theme(title = element_text(face = "bold"),
                axis.title = element_text(face = "bold"))

```

# Model

Let's fit a linear model to get the slope (b) and intercept(a).

$$
y = a + bx
$$

```{r}
fl_mod <- lm(fluo ~ conc_pgml, data = data)
summary(fl_mod)
```

From above, we can see that slope = 1.9304, and intercept = 1.5179.

# Errors in slope and intercept of regression line

The limit of detection is defined as:

$$
LOD = \gamma_B + 3_{SB}
$$
where LOD is the analyte concentration wich gives a signal equal to the blank signal plus three standard deviations of the blank.

A function was created to calculate LOD and LOQ:

```{r, echo = TRUE}
calcLOD_y <- function(model) {
  SSE <- sum(model$residuals**2)
  n <- length(model$residuals) -2
  Syx <- sqrt(SSE/n)
  intercept <- as.numeric(model$coefficients[1])
  calculated_y <- intercept + 3*Syx
  names(calculated_y) <- "calculated_y"
  print(calculated_y)
  
  chemCal::inverse.predict(model,
                  newdata = calculated_y,
                  alpha = 0.05) 
}
```
```{r, echo = TRUE}
calcLOQ_y <- function(model) {
  SSE <- sum(model$residuals**2)
  n <- length(model$residuals) -2
  Syx <- sqrt(SSE/n)
  intercept <- as.numeric(model$coefficients[1])
  calculated_y <- intercept + 10*Syx
  names(calculated_y) <- "calculated_y"
  print(calculated_y)
  
  chemCal::inverse.predict(model,
                  newdata = calculated_y,
                  alpha = 0.05) 
}
```

Inserting the linear model from the fluorescence data:

```{r, echo = TRUE}
LOD_x <- calcLOD_y(fl_mod)
LOD_x$Prediction 
```

```{r, echo = TRUE}
LOQ_x <- calcLOQ_y(fl_mod)
LOQ_x$Prediction 
```

# Predict the value of x from y:

To predict the concentration of fluorescein that has fluorescence units of 2.9, we use the function **inverse.predict()**:

```{r, echo = TRUE}
chemCal::inverse.predict(fl_mod, 
                newdata = 2.9,
                alpha = 0.05)
```
